Raw Data Comparison

### 2019-09-10 Raw Data Comparison.Rmd
...

---
title: "Raw Data Comparison"
author: "Sarah Killcoyne"
date: "23/08/2019"
output: 
  html_document: 
    fig_width: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(BarrettsProgressionRisk)
library(reshape2)

@@

```{r, fig.width=12, fig.height=12}
# par(mfrow=c(4,2))
# qqnorm( na.omit(tfit.data$D701_D501_10722_11823), main='Training sample' ) 
# qqnorm( na.omit(vfit.data$`SLX-17976.N705-S508`), main='Validation sample' )
# 
# qqnorm( na.omit(tfit.data$D701_D505_10722_11823), main='Training sample' ) 
# qqnorm( na.omit(vfit.data$`SLX-17977.N706-S510`), main='Validation sample' )
# 
# qqnorm(na.omit(tfit.data$D706_D502_13696), main='Training sample')
# qqnorm(na.omit(vfit.data$`SLX-16758.N703-S511`), main='Validation sample') # this was one of the OAC samples
# 
# qqnorm(na.omit(tfit.data$D712_D504_13692), main='Training sample') # P sample, correctly predicted
# qqnorm(na.omit(vfit.data$`SLX-16783.N715-S505`), main='Validation sample') # P sample, correctly predicted

@@

# x %>% dplyr::filter(mean >= 1.05)
# x %>% dplyr::filter(sd >= 0.2) 
# 
# x %>% dplyr::filter(var > 0.05)
# x %>% dplyr::filter(IQR >= 0.2)

@@

```{r, echo=F, message=F, warning=F}
files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', '5e06_cleaned_tiled', recursive=T, full.names = T)
training.tiles = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% rename(X1 = 'Sample')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', '5e06.*MSE', recursive=T, full.names = T)
training.resids = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% rename(X1 = 'Sample')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/', '5e06_tiles', recursive=T, full.names = T)
val.tiles = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% rename(X1 = 'Sample')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/', '5e06_MSE', recursive=T, full.names = T)
val.resids = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% rename(X1 = 'Sample')

@@

Per patient after tiling which merges the segments using a weighted mean.  Again, there's an upwards shift per-patient. 
```{r, fig.width=10}
per.pt.tile<-function(df) {
  means = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~mean(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='mean')
  sds = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~sd(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='sd')
  vars = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~var(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='var')
  iqrs = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~IQR(.,na.rm=T))) %>% t %>% as_tibble %>% rename(V1='IQR')

@@

```{r}
tile.dist.measure<-function(df) {
  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='mean') 
  sds = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='mean') 
  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~sd(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='sd') 
  iqrs = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~IQR(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='IQR') 
  vars = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~var(.,na.rm=T))) %>% t, rownames='bin')  %>% rename(V1='var')

  dist = means %>% full_join(sds, by='bin') %>% full_join(iqrs, by='bin') %>% full_join(vars, by='bin')
  return(dist)
}

@@

ggplot(full_join(t.dist, v.dist, by='bin'), aes(mean.x,mean.y)) + geom_point() + labs(x='train', y='val', title='mean')

ggplot(full_join(t.dist, v.dist, by='bin'), aes(sd.x,sd.y)) + geom_point() + labs(x='train', y='val', title='SD')

#dens.obs = density(training.tiles$`1:1-4985012`)

#resample.obs = sample(dens.obs$x, nrow(val.tiles), replace=TRUE, prob=dens.obs$y)

@@

After mean normalization per-bin using the z scores from the training data. And sure enough, it spreads the values out considerably for the validation data.  So the question is if the validation data were adjusted would it do this?

@@

But what if I, as a quick test, normalize the validation cohort with itself?

v.means = apply(as.matrix(val.tiles[,-1]),2, mean, na.rm=T)
v.sd = apply(as.matrix(val.tiles[,-1]),2, sd, na.rm=T)

@@

@@

find.rank<-function(r,x) {
  if ( nrow(x %>% filter(rank == r)) > 0 ) return(x %>% filter(rank == r))
  find.rank(r-1,x)
  # if (nrow(x) <= 4) return(x)
  # (median(x$rank))
  # nrow(x)
  # if (r <= floor(median(x$rank)) ) {
  #   find.rank(r, x %>% filter(rank <= median(rank)))
  # } else {
  #   find.rank(r, x %>% filter(rank >= median(rank)))
  # }
}

adj.v = val.tiles
for (n in 2:ncol(training.tiles)) {
  print(n)
  col = colnames(training.tiles)[n]

@@

j = 100
hist(adj.v[[j]], breaks=20, xlim=c(0.84,1.1))
hist(val.tiles[[j]], breaks=20, xlim=c(0.84,1.1))
hist(training.tiles[[j]], breaks=20, xlim=c(0.84,1.1))

@@

## Adjusted by directly altering validation values for the nearest training value?

Just a quick try, but could directly altering the values of the validation data to the nearest training value alter these?  The answer appears to be 'no'.

```{r, eval=F}

adj.val.tiles = val.norm
for (i in 1:nrow(adj.val.tiles)) {
#  print(i)
  for (j in 2:ncol(adj.val.tiles[i,])) {
    bin = tt.norm[,j,drop=T]
    bq = quantile(bin, probs=seq(0,1,0.2))

    iqr = range(quantile(bin,0.25), quantile(bin,0.75))

    if (adj.val.tiles[i,j,drop=T] >= min(iqr) & adj.val.tiles[i,j,drop=T] <= max(iqr)) {
      mn = min(iqr); mx = max(iqr)
    } else {
      mn = max(bq[which(bq <= adj.val.tiles[i,j,drop=T] )])
      mx = min(bq[which(bq >= adj.val.tiles[i,j,drop=T] )])
    }

    adj.val.tiles[i,j] = sample(bin[bin >= mn & bin <= mx],1)

    #adj.val.tiles[i,j] = bin[which.min((bin-adj.val.tiles[i,j,drop=T])^2)]
  }
}

hist(tt.norm$`1:1-4985012`, breaks=20)
hist(adj.val.tiles$`1:1-4985012`, breaks=20)
hist(val.norm$`1:1-4985012`, breaks=20)

adj.v.pt = per.pt.tile(adj.val.tiles) %>% mutate(cohort='adj.val')

ggplot(bind_rows(t.pt, v.pt, adj.v.pt), aes(cohort, mean, group=cohort, fill=cohort)) + 
  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
  labs(title='Mean per patient')

v.adj.dist = tile.dist.measure(adj.val.tiles) %>% mutate(cohort='adj val')

ggplot(bind_rows(t.dist, v.dist,v.adj.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
  labs(title='Mean per bin')

adj.val.norm = as.matrix(adj.val.tiles[-1])
for (i in 1:ncol(adj.val.norm))
  adj.val.norm[,i] = BarrettsProgressionRisk:::unit.var(adj.val.norm[,i], z.mean[i], z.sd[i])

adj.val.norm = as_tibble(adj.val.norm) %>% add_column('Sample'=adj.val.tiles$Sample,.before=T)
vadj.norm.dist = tile.dist.measure(adj.val.norm) %>% mutate(cohort='adj val')

ggplot(bind_rows(tnorm.dist, vnorm.dist,vadj.norm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
  labs(title='Mean per bin after norm')
```

## Normalize per patient instead?

```{r}
t.pt
v.pt

median(v.pt$var)

median(t.pt$var)
median(t.pt$sd)
median(t.pt$mean)

summary(unlist(val.tiles[1,-1]))
var(unlist(val.tiles[1,-1]))

fourier<-function(x) {
  fx = fft(x)
  plot(Re(fx), type='l', xlim=c(0,20))
  fx[20:length(fx)] = 0+0i
  fxx = fft(fx, inverse = TRUE)/length(fx)

#  layout(matrix(1:2,2,1)) 
#  plot(x, type="l", main="Original Data")
#  plot(Re(fxx),type="l", main="Fourier Transform Filtering")

#  hist(x, breaks=50)
#  hist(Re(fxx), breaks=50)
  Re(fxx) 
}

fourier.tiles = val.tiles

for (i in 1:nrow(fourier.tiles))
  fourier.tiles[i,-1] = fourier(unlist(fourier.tiles[i,-1]))

ft.disp = per.pt.tile(fourier.tiles) %>% mutate(cohort='fourier transform')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.pt, v.pt, ft.disp), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean per patient'),
  ggplot(bind_rows(t.pt, v.pt, ft.disp), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='StdDev per patient'),
  ncol=2, top='Per patient tiled values')

f.dist = tile.dist.measure(fourier.tiles) %>% mutate(cohort='fourier transform')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.dist, v.dist, f.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(t.dist, v.dist, f.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'), 
ncol=2, top='Per bin tiled values')

fourier.norm = as.matrix(fourier.tiles[-1])
for (i in 1:ncol(fourier.norm))
  fourier.norm[,i] = BarrettsProgressionRisk:::unit.var(fourier.norm[,i], z.mean[i], z.sd[i])

ft.norm = as_tibble(fourier.norm) %>% add_column('Sample'=fourier.tiles$Sample,.before=T)

ftnorm.dist = tile.dist.measure(ft.norm) %>% mutate(cohort='fourier')

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vnorm.dist, ftnorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vnorm.dist, ftnorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin normalized tile values')

@@

```{r cosigndiff, eval=F}
## What about finding the nearest patient via cosine diff?

cos.sim=function(ma, mb){
  mat=tcrossprod(ma, mb)
  t1=sqrt(apply(ma, 1, crossprod))
  t2=sqrt(apply(mb, 1, crossprod))
  mat / outer(t1,t2)
}

cs.pt = do.call(rbind, lapply(1:nrow(val.tiles), function(j) {
  sapply(1:nrow(training.tiles), function(i) {
    cos.sim(as.matrix(val.tiles[j,-1]), as.matrix(training.tiles[i,-1]))
  })
}))

@@

>>>>

---
title: "Raw Data Comparison"
author: "Sarah Killcoyne"
date: "23/08/2019"
output: 
  html_document: 
    fig_height: 4
    fig_width: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(BarrettsProgressionRisk)
library(reshape2)
library(gridExtra)
library(e1071)

@@

```{r, fig.width=12, fig.height=12}

@@

@@

```{r, echo=F, message=F, warning=F}
files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', '5e06_cleaned_tiled', recursive=T, full.names = T)
training.tiles = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', '5e06.*MSE', recursive=T, full.names = T)
training.resids = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/', '5e06_tiles', recursive=T, full.names = T)
val.tiles = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/', '5e06_MSE', recursive=T, full.names = T)
val.resids = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

@@

Per patient after tiling which merges the segments using a weighted mean.  Again, there's an upwards shift per-patient. 
```{r, fig.width=10}
per.pt.tile<-function(df) {
  means = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~mean(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('mean' = 'V1')
  sds = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~sd(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('sd' = 'V1')
  vars = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~var(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('var' = 'V1')
  iqrs = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~IQR(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('IQR' = 'V1')

@@

```{r}
tile.dist.measure<-function(df) {
  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('mean' = 'V1')
  sds = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~sd(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('sd' = 'V1')
  iqrs = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~IQR(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('iqrs' = 'V1')
  vars = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~var(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('var' = 'V1')

  skew = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~skewness(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('skew' = 'V1')
  kurt = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~kurtosis(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('kurtosis' = 'V1')

  dist = means %>% full_join(sds, by='bin') %>% full_join(iqrs, by='bin') %>% full_join(vars, by='bin') %>% full_join(skew, by='bin') %>% full_join(kurt, by='bin')
  return(dist)
}

@@

gridExtra::grid.arrange(
  ggplot(bind_rows(t.dist, v.dist), aes(cohort, skew, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Skew'),
  ggplot(bind_rows(t.dist, v.dist), aes(cohort, kurtosis, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Kurtosis'), 
  ncol=2, top='Per bin tiled values')

grid.arrange(
  ggplot(full_join(t.dist, v.dist, by='bin'), aes(mean.x,mean.y)) + geom_point() + labs(x='train', y='val', title='mean'),
  ggplot(full_join(t.dist, v.dist, by='bin'), aes(sd.x,sd.y)) + geom_point() + labs(x='train', y='val', title='SD'),
  ggplot(full_join(t.dist, v.dist, by='bin'), aes(var.x,var.y)) + geom_point() + labs(x='train', y='val', title='var'),
  ggplot(full_join(t.dist, v.dist, by='bin'), aes(skew.x,skew.y)) + geom_point() + labs(x='train', y='val', title='Skew'),
ncol=2)

@@

After scaling and centering per-bin using the z scores from the training data. And sure enough, it spreads the values out considerably for the validation data.  So the question is if the validation data were adjusted would it do this?

@@

But what if I, as a quick test, normalize the validation cohort with itself? This looks better here, but when I use this for predictions it doesn't significantly change the results.

```{r}
v.means = apply(as.matrix(val.tiles[,-1]),2, mean, na.rm=T)
v.sd = apply(as.matrix(val.tiles[,-1]),2, sd, na.rm=T)

@@

This moved the non-progressors the right direction with the predictions, but there was still a 30% FPR for non-progressors.  The age-bias reappeared for the progressors though.

@@

find.rank<-function(r,x) {
  if ( nrow(x %>% filter(rank == r)) > 0 ) return(x %>% filter(rank == r))
  find.rank(r-1,x)
}

adj.v = val.tiles
for (n in 2:ncol(training.tiles)) {
  col = colnames(training.tiles)[n]

@@

@@

### Predictions

```{r}
preds = do.call(bind_rows, purrr::map(
  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions-rankadj/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`)

preds = left_join(all.val, preds, by=c('Samplename'='Sample', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))

# Non progressors
preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n)

# FPR is nearly 40% still
nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP'))

# Progressors
preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n)

# FNR is still over 50% in the old progressors as well
nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`)))
```

## Distributions

How does it change if I use a global variance, or a rolling-bin variance (to preserve some local info) to adjust?

```{r}
# global variance
all.tiles = bind_rows(training.tiles, val.tiles)

means = apply(all.tiles[,-1], 2, mean)
sds = apply(all.tiles[,-1], 2, sd)
global.sd = sd(as.matrix(all.tiles[,-1]))

wt.sds = sapply(sds, function(x) weighted.mean(x,global.sd,w=c(0.2)))

plot(wt.sds, sds)

tt.norm.gv = as.matrix(training.tiles[-1])
for (i in 1:ncol(tt.norm.gv))
  tt.norm.gv[,i] = BarrettsProgressionRisk:::unit.var(tt.norm.gv[,i], means[i], wt.sds[i])

val.norm.gv = as.matrix(val.tiles[-1])
for (i in 1:ncol(tt.norm.gv))
  val.norm.gv[,i] = BarrettsProgressionRisk:::unit.var(val.norm.gv[,i], means[i], wt.sds[i])

all.norm.gv = as.matrix(all.tiles[-1])
for (i in 1:ncol(all.norm.gv))
  all.norm.gv[,i] = BarrettsProgressionRisk:::unit.var(all.norm.gv[,i], means[i], wt.sds[i])

tt.norm.gv = as_tibble(tt.norm.gv) %>% add_column('Sample'=training.tiles$Sample,.before=T)
val.norm.gv = as_tibble(val.norm.gv) %>% add_column('Sample'=val.tiles$Sample,.before=T)
all.norm.gv = as_tibble(all.norm.gv) %>% add_column('Sample'=all.tiles$Sample,.before=T)

tt.norm.dist = tile.dist.measure(tt.norm.gv) %>% mutate(cohort='train, wt var')
vp.norm.dist = tile.dist.measure(val.norm.gv) %>% mutate(cohort='val, wt var')
all.norm.dist = tile.dist.measure(val.norm.gv) %>% mutate(cohort='all, wt var')

gridExtra::grid.arrange(
  ggplot(bind_rows(tt.norm.dist, vp.norm.dist, all.norm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tt.norm.dist, vp.norm.dist, all.norm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin wt var scale values')

@@

tx = melt(training.tiles)
vx = melt(val.tiles)

grid.arrange(ggplot(tx, aes(value)) + xlim(0.4, 1.5) + geom_histogram(bins=50, color='lightblue') + labs(title='Training'),
             ggplot(vx, aes(value)) + xlim(0.4,1.5) + geom_histogram(bins=50, color='lightblue') + labs(title='Validation'))

# The extreme values in the validation set come from the OACs
all.val %>% filter(Samplename %in% unique(vx[which(vx$value > 1.5),'Sample']))

vx = vx %>% filter(Sample %in% (all.val %>% filter( !grepl('OCCAMS',`Hospital Research ID`)) %>% dplyr::select(Samplename) %>% pull) )

tx.normal = fitdistrplus::fitdist(tx$value, 'norm')
vx.normal = fitdistrplus::fitdist(vx$value, 'norm')

@@

From training data
```{r}
plot(tx.normal)
```

From validation data
```{r}
plot(vx.normal)
```

train.P = round(pnorm(tx$value,mean=tx.normal$estimate['mean'], sd=tx.normal$estimate['sd']),4)
val.P = round(pnorm(as.matrix(val.tiles[-1]),mean=tx.normal$estimate['mean'], sd=tx.normal$estimate['sd']),4)
#val.P = round(pnorm(as.matrix(val.tiles[-1]),mean=vx.normal$estimate['mean'], sd=vx.normal$estimate['sd']),4)

p.transform = as.matrix(val.tiles[,-1])
for (i in 1:nrow(p.transform)) {
  for (j in 1:ncol(p.transform)) {
    p.transform[i,j] = mean(tx[which(train.P == val.P[i,j]),'value'])
  }
}

n = 20
hist(as.matrix(training.tiles[,-1])[,n] ,breaks=20)
hist(as.matrix(val.tiles[,-1])[,n] ,breaks=20)
hist(p.transform[,n], breaks=20)

# so now if I scale...

vp.norm = p.transform
for (i in 1:ncol(vp.norm))
  vp.norm[,i] = BarrettsProgressionRisk:::unit.var(vp.norm[,i], means[i], sd[i])
vp.norm = as_tibble(vp.norm) %>% add_column('Sample'=val.tiles$Sample,.before=T)

vp.norm.dist = tile.dist.measure(vp.norm) %>% mutate(cohort='val p')

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vnorm.dist, vp.norm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vnorm.dist,vp.norm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin normalized tile values')
